---
title: "make non-trait community metrics"
author: "Tim Ohlert"
date: "12/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### This code is meant to calculate non-trait metrics for all Sevilleta control plots for all time. 



# Add libraries


```{r, message = FALSE}
library(plyr)
library(lmerTest)
library(tidyverse)
library(viridis)
library(MuMIn)
library(ggeffects)
library(ggthemes)

```


# load data and clean

## load plant community data

```{r}
corebiomass <- read.csv("~/Misc Grad school/Sevilleta_allbiomass_22Jan2020.csv")

corebiomass <- subset(corebiomass, 
                      site != "crust_creosote" &
                      site != "crust_grass" &
                      site != "crust_PJ" &
                      site != "grid_black" &
                      site != "grid_blue" &
                      site != "grid_creosote" &
                      site != "iso_web" &
                      site != "tower_west"
                      )


```

## reduce data to fall controls and summarize including calculating percent abundance of species within communities

```{r}
core_5 <- subset(corebiomass,
                 treatment == "C" &
                   season == "fall"
                 )

core_3 <- ddply(core_5, c("site", "year", "web", "transect", "block", "plot", "subplot", "quad", "SPEI.comp"),
                function(x)data.frame(
                  sr = length(x$kartez)
                ))


#core_2 <- merge(core_3, core_5, by = c("site", "year", "web", "transect", "block", "plot", "subplot", "quad"))
#core_2$percent.abundance <- core_2$cover/core_2$plot.cover

#calculate number of species in each plot 
#core_obs_number <- ddply(core_2, c("site", "year", "web", "transect", "block", "plot", "subplot", "quad"),
#                         function(x)data.frame(
#                           obs.number = as.numeric(length(x$kartez))
#                         ))

#remove plots with only a few species because otherwise the hypervolume doesn't work
#core_1 <- merge(core_2, core_obs_number, by = c("site", "year", "web", "transect", "block", "plot", "subplot", "quad"))
#core <- subset(core_1, obs.number > 3)
```


# let's graph some shit

##models
```{r}
srmod.1 <- lmer(sr ~ SPEI.comp + (1|year/site), data = core_3)
srmod.2 <- lmer(sr ~ SPEI.comp + I(SPEI.comp^2) + (1|year/site), data = core_3)
srmod.3 <- lmer(sr ~ SPEI.comp  + I(SPEI.comp^2) + I(SPEI.comp^3) + (1|year/site), data = core_3)
srmod.4 <- lmer(sr ~ SPEI.comp + I(SPEI.comp^2) + I(SPEI.comp^3) + I(SPEI.comp^4) + (1|year/site), data = core_3)
srmod.5 <- lmer(sr ~ SPEI.comp + I(SPEI.comp^2) + I(SPEI.comp^3) + I(SPEI.comp^4) + I(SPEI.comp^5) + (1|year/site), data = core_3)

AICc(srmod.1, srmod.2, srmod.3, srmod.4, srmod.5)

```

```{r}
summary(srmod.1)
r.squaredGLMM(srmod.1)
```



## graph
```{r}
x <- ggpredict(srmod.1, c("SPEI.comp"))
plot(x, ci=FALSE)+
  geom_ribbon(aes(ymin=predicted-std.error,ymax=predicted+std.error), alpha =.5)+
  geom_point(data=core_3, aes(x=SPEI.comp,y=sr, colour = site))+
  xlab("SPEI")+
  ylab("Species richness")+
  ggtitle("")+
  scale_color_viridis(discrete = TRUE, option = "D")+
  theme_base()
```




```{r}
model.list <- lme4::lmList(sr~SPEI.comp  | site, core_3)
```

```{r}
summary(model.list)
```

```{r}

df <-data.frame(change = c("No_change", "Positive", "Negative"),
                sites = c(9,1,3))
                          
df$change <- factor(df$change, levels = c("No_change","Positive","Negative"))
                           

ggplot(df, aes(change, sites))+
  geom_bar(stat="identity")+
  ylim(0,13)+
  ylab("# of sites")+
  xlab("")+
  theme_base()


```

```{r}
ggplot(core_3, aes(SPEI.comp,sr))+
  facet_wrap(~site)+
  geom_smooth(method = "lm")+
  geom_point(data=core_3, aes(x=SPEI.comp,y=sr, colour = site))+
  xlab("SPEI")+
  ylab("Species richness")+
  ggtitle("")+
  scale_color_viridis(discrete = TRUE, option = "D")+
   theme(legend.position = "none")#+
  #theme_base()
```




# change between years

                       
## make delta metrics
```{r}

#make replicate column

core_3 <- unite(core_3, replicate, c("site","web","plot","quad","transect","plot","subplot"), sep=".",remove=FALSE)

dummy <- core_3

dummy$year <- core_3$year - 1

dummy <- dummy %>% 
  rename(
    sr.2 = sr,
    SPEI.comp.2 = SPEI.comp
        )


dummy <- dummy[,c("replicate","year", "sr.2", "SPEI.comp.2")]

core_3_prevyear <- merge(core_3, dummy, by = c("replicate", "year"))

core_3_prevyear$sr_delta <- core_3_prevyear$sr - core_3_prevyear$sr.2
core_3_prevyear$SPEI_delta <- core_3_prevyear$SPEI.comp - core_3_prevyear$SPEI.comp.2


core_3_delta <- core_3_prevyear[,c("replicate", "year", "site", "sr_delta", "SPEI_delta")]


```



## model alpha change
```{r}
srmod.1 <- lmer(sr_delta ~ SPEI_delta + (1|year/site), data = core_3_delta)
srmod.2 <- lmer(sr_delta ~ SPEI_delta + I(SPEI_delta^2) + (1|year/site), data = core_3_delta)
srmod.3 <- lmer(sr_delta ~ SPEI_delta  + I(SPEI_delta^2) + I(SPEI_delta^3) + (1|year/site), data = core_3_delta)
srmod.4 <- lmer(sr_delta ~ SPEI_delta + I(SPEI_delta^2) + I(SPEI_delta^3) + I(SPEI_delta^4) + (1|year/site), data = core_3_delta)
srmod.5 <- lmer(sr_delta ~ SPEI_delta + I(SPEI_delta^2) + I(SPEI_delta^3) + I(SPEI_delta^4) + I(SPEI_delta^5) + (1|year/site), data = core_3_delta)

AICc(srmod.1, srmod.2, srmod.3, srmod.4, srmod.5)

```

```{r}
summary(srmod.1)
r.squaredGLMM(srmod.1)
```



## graph alpha change

```{r}
x <- ggpredict(srmod.1, c("SPEI_delta"))
plot(x, ci=FALSE)+
  geom_ribbon(aes(ymin=predicted-std.error,ymax=predicted+std.error), alpha =.5)+
  geom_point(data=core_3_delta, aes(x=SPEI_delta,y=sr_delta, colour = site))+
  xlab("Change in SPEI")+
  ylab("Change in species richness")+
  ggtitle("")+
  scale_color_viridis(discrete = TRUE, option = "D")+
  geom_hline(yintercept = 0)+
  geom_vline(xintercept = 0)+
  theme_base()
```






```{r}
model.list <- lme4::lmList(sr_delta~SPEI_delta  | site, core_3_delta)
```

```{r}
summary(model.list)
```

```{r}

df <-data.frame(change = c("No_change", "Positive", "Negative"),
                sites = c(4,1,8))
                          
df$change <- factor(df$change, levels = c("No_change","Positive","Negative"))
                           

ggplot(df, aes(change, sites))+
  geom_bar(stat="identity")+
  ylim(0,13)+
  ylab("# of sites")+
  xlab("")+
  theme_base()


```

```{r}
ggplot(core_3_delta, aes(SPEI_delta,sr_delta))+
  facet_wrap(~site)+
  geom_smooth(method = "lm")+
  geom_point(data=core_3_delta, aes(x=SPEI_delta,y=sr_delta, colour = site))+
  xlab("Change in SPEI")+
  ylab("Change in species richness")+
  ggtitle("")+
  scale_color_viridis(discrete = TRUE, option = "D")+
  geom_hline(yintercept = 0)+
  geom_vline(xintercept = 0)+
   theme(legend.position = "none")#+
  #theme_base()
```



























